European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-16 manuscript/manuscript.qmd
Adjust spelling for last para
Robustness: Continental Preference? (#33)
* add europe-disaggregated table

* update  regtable_europe for new tidytable version

* Add robustness draft

* use `\times` rather than x

* Fix variable line break

* Add question mark to robustness title

* Rearrange paragraphs to add Greece example

* move interpretation below the table

* Fix mismatched divider in table

* Add FE and regressiopn type to table

* Fix table captions

* update interaction term interpretation

* Add european recipient coefficient interpretation

* Adjust spelling for last para

-   IMF chief European [@kahler2001]

@@

These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF. In other words, European powers appear to exercise their influence globally, rather than selectively favoring their regional neighbours.

>>>>

As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance. However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay. This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994]. Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process. For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts. Specifically, this applies to countries that align with the U.S. in United Nations General Assembly and United Nations Security Council voting [@lipscy2018; @thacker1999; @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004]. This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2002; @woods2003; @clark2020]. The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.

The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons. Existing studies provide substantial evidence of U.S. influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S. and other major powers, particularly the European countries. IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous. The organization is traditionally chaired by an European national [@kahler2001],

@@

These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF. In other words, European powers appear to exercise their influence globally, rather than selectively favouring their regional neighbours.
